---
title: "demo-exports"
output: html_document
date: "2025-03-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## read data

```{r cars}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(plotly)
library(tidyr)
library(countrycode)
exports_raw <- readRDS("2021_exports_raw.rds")
```

## monthly overview exports

```{r}
monthly_exports_agg <- function(data, ...) {
  data %>%
    group_by(...) %>%
    filter((exports_nzd_fob != 0) | (reexports_nzd_fob != 0)) %>%
    summarise(
      Total_Only_Exports_nzd_fob = sum(exports_nzd_fob, na.rm = T),
      Total_Only_Exports_qty = sum(exports_qty, na.rm = T),           
      Total_Reexports_nzd_fob = sum(reexports_nzd_fob, na.rm = T),
      Total_Reexports_qty = sum(reexports_qty, na.rm = T),        
      Total_Exports_nzd_fob = sum(total_exports_nzd_fob, na.rm = T),  
      Total_Exports_qty = sum(total_exports_qty, na.rm = T)        
    ) %>%
    mutate(
      unit_fob = Total_Exports_nzd_fob / Total_Exports_qty,  
      unit_reexports_fob = Total_Reexports_nzd_fob / Total_Reexports_qty, 
      unit_only_exports_fob = Total_Only_Exports_nzd_fob / Total_Only_Exports_qty, 
      export_qty_proportion = Total_Only_Exports_nzd_fob / Total_Exports_nzd_fob, 
    )
}
overview_plot <- function(data, ...){
  agg_data <- data %>%
    monthly_exports_agg(...)

  p1 <- ggplot(agg_data) +
    geom_line(aes(x = month, y = Total_Only_Exports_nzd_fob / 1e6, color = "Only_Exports"),  size = 1) +
    geom_line(aes(x = month, y = Total_Reexports_nzd_fob / 1e6, color = "Reexports"), size = 1) +
    geom_line(aes(x = month, y = Total_Exports_nzd_fob / 1e6, color = "Exports"), size = 1) +
    labs(title = "Monthly Imports fob($NZD)", 
         x = "Month", 
         y = "Total Imports ($NZD, in millions)") +
    theme_minimal() +
    scale_color_manual(values = c("blue", "red", "green"))

  p2 <- ggplot(agg_data) +
    geom_line(aes(x = month, y = Total_Only_Exports_qty/1e7, color = "Total Only Exports Qty"), size = 1,) + 
    geom_line(aes(x = month, y = Total_Reexports_qty/1e7, color = "Total Reexports Qty"), size = 1,) +
    geom_line(aes(x = month, y = Total_Exports_qty/1e7, color = "Total Exports Qty"), size = 1,) +
    labs(title = "Monthly Imports Total Imports Qty", 
         x = "Month", 
         y = "Total Imports Qty (in ten millions)") +
    theme_minimal()
  
  p3 <- ggplot(agg_data) +
    geom_line(aes(x = month, y = unit_fob, color = "unit_fob"), size = 1,) + 
    geom_line(aes(x = month, y = unit_reexports_fob, color = "unit_reexports_fob"), size = 1,) +
    geom_line(aes(x = month, y = unit_only_exports_fob, color = "unit_only_exports_fob"), size = 1,) +
   labs(title = "Monthly Imports unit_fob", 
         x = "Month", 
         y = "unit") +
    theme_minimal() + 
    scale_color_manual(values = c("blue", "red", "green"))
  
  print(p1)
  print(p2)
  print(p3)
}

exports_raw %>% overview_plot(month)
```
Most of them are exported directly, and a small part is re-exported, but the unit price of re-export is relatively high.

易于比较，需要聚焦到一个小的部分

## global geographical map

```{r}
agg_data <- exports_raw %>%
    group_by(iso3) %>%
    filter(exports_nzd_fob != 0 | reexports_nzd_fob !=0) %>%
    summarise(
      Total_Only_Exports_nzd_fob = sum(exports_nzd_fob, na.rm = T),
      Total_Only_Exports_qty = sum(exports_qty, na.rm = T),           
      Total_Reexports_nzd_fob = sum(reexports_nzd_fob, na.rm = T), 
      Total_Reexports_qty = sum(reexports_qty, na.rm = T),       
      Total_Exports_nzd_fob = sum(total_exports_nzd_fob, na.rm = T), 
      Total_Exports_qty = sum(total_exports_qty, na.rm = T)  
    )

plot_geo(agg_data) %>%
  add_trace(
    z = ~Total_Only_Exports_nzd_fob,
    locations = ~iso3,
    colorscale = "Viridis"
  ) %>%
  layout(title = "Total Only Exports fob ($NZD)")


plot_geo(agg_data) %>%
  add_trace(
    z = ~Total_Reexports_nzd_fob,
    locations = ~iso3,
    colorscale = "Viridis"
  ) %>%
  layout(title = "Total Reexport fob ($NZD)")


plot_geo(agg_data) %>%
  add_trace(
    z = ~Total_Exports_nzd_fob,
    locations = ~iso3,
    colorscale = "Viridis"
  ) %>%
  layout(title = "Total Exports fob ($NZD)")
```

qty 不能比较，需要同类型的 ，literture review，-> concrete

For re-export, it is mainly Australia and USA, and for direct export, it is mainly China, Australia and USA.

## group_by_continent

```{r}
library(gridExtra)
library(grid)
library(cowplot)

plotv2 <- function(data,show_group,...,filter_conditions = list()){
  
  agg_data <- data %>%
    monthly_exports_agg(...)
  
  if (length(filter_conditions) > 0) {
    for (filter_cond in filter_conditions) {
      agg_data <- agg_data %>% filter(!!parse_expr(filter_cond))  # Apply each filter condition
    }
  }
  
  p1 <- ggplot(agg_data) +
    geom_line(aes(x = month, y = Total_Only_Exports_nzd_fob / 1e6, color = {{show_group}}),  size = 1) +
    labs(title = paste("Monthly Only Exports ($NZD fob) group by",substitute(show_group)),
         x = "Month",
         y = "Total Exports fob ($NZD, in millions)") +
    theme_minimal()
  
  p2 <- ggplot(agg_data) +
    geom_line(aes(x = month, y = Total_Only_Exports_qty/1e7, color = {{show_group}}), size = 1,) + 
    labs(title = paste("Monthly Only Exports Qty group by",substitute(show_group)),
         x = "Month",
         y = "Total Exports Qty (in ten millions)") +
    theme_minimal()
  
  p3 <- ggplot(agg_data) +
    geom_line(aes(x = month, y = unit_fob, color = {{show_group}}), size = 1,) + 
    labs(title = paste("Monthly unit_fob group by",substitute(show_group)),
         x = "Month",
         y = "unit") +
    theme_minimal()
  
  p4 <- ggplot(agg_data) +
    geom_line(aes(x = month, y = export_qty_proportion, color = {{show_group}}), size = 1,) + 
    labs(title = paste("Monthly Exports Proportion group by",substitute(show_group)),
         x = "Month",
         y = "Export Quantity Proportion") +
    theme_minimal()
   
  p5 <- ggplot(agg_data) +
    geom_line(aes(x = month, y = Total_Reexports_nzd_fob / 1e6, color = {{show_group}}),  size = 1) +
    labs(title = paste("Monthly Reexports ($NZD fob) group by",substitute(show_group)),
         x = "Month",
         y = "Total Reexports fob ($NZD, in millions)") +
    theme_minimal()
  
  p6 <- ggplot(agg_data) +
    geom_line(aes(x = month, y = Total_Reexports_qty/1e7, color = {{show_group}}), size = 1,) + 
    labs(title = paste("Monthly Reexports Qty group by",substitute(show_group)),
         x = "Month",
         y = "Total Reexports Qty (in ten millions)") +
    theme_minimal()
  
  print(p1)
  print(p2)
  print(p3)
  print(p4)
  print(p5)
  print(p6)
  agg_data
}

pie_plot <- function(agg_data,show_group){
  agg_data_pie <- agg_data %>%
    group_by({{show_group}}) %>%
    summarise(Total_only_exports_fob = sum(Total_Only_Exports_nzd_fob, na.rm = T),
              Total_only_exports_qty = sum(Total_Only_Exports_qty, na.rm = T),
              Total_reexports_nzd_fob = sum(Total_Reexports_nzd_fob, na.rm = T),
              Total_reexports_qty = sum(Total_Reexports_qty, na.rm = T)) %>%
    mutate(Percentage_only_exports_fob = Total_only_exports_fob / sum(Total_only_exports_fob) * 100,
          Percentage_only_exports_qty = Total_only_exports_qty / sum(Total_only_exports_qty) * 100,
          Percentage_reexports_fob = Total_reexports_nzd_fob / sum(Total_reexports_nzd_fob) * 100,
          Percentage_reexports_qty = Total_reexports_qty / sum(Total_reexports_qty) * 100)
  
  # 画饼图
  p1 <- ggplot(agg_data_pie, aes(x = "", y = Percentage_only_exports_fob, fill = {{show_group}})) +
    geom_bar(width = 1, stat = "identity") +
    coord_polar("y", start = 0) +
    # labs(title = paste("Percentage of Total Only Exports fob group by", substitute(show_group)),
    #      x = NULL,
    #      y = NULL) +
    theme_void() +
    theme(legend.position = "none")
  
  p2 <- ggplot(agg_data_pie, aes(x = "", y = Percentage_only_exports_qty, fill = {{show_group}})) +
    geom_bar(width = 1, stat = "identity") +
    coord_polar("y", start = 0) +
    # labs(title = paste("Percentage of Total Only Exports qty group by", substitute(show_group)),
    #      x = NULL,
    #      y = NULL) +
    theme_void() +
    theme(legend.position = "none")
  
  p3 <- ggplot(agg_data_pie, aes(x = "", y = Percentage_reexports_fob, fill = {{show_group}})) +
    geom_bar(width = 1, stat = "identity") +
    coord_polar("y", start = 0) +
    # labs(title = paste("Percentage of Total Reexports fob group by", substitute(show_group)),
    #      x = NULL,
    #      y = NULL) +
    theme_void() +
    theme(legend.position = "none")
  
  p4 <- ggplot(agg_data_pie, aes(x = "", y = Total_reexports_qty, fill = {{show_group}})) +
    geom_bar(width = 1, stat = "identity") +
    coord_polar("y", start = 0) +
    # labs(title = paste("Percentage of Total Reexports qty group by", substitute(show_group)),
    #      x = NULL,
    #      y = NULL) +
    theme_void() +
    theme(legend.position = "none")
  
  legend <- get_legend(
    ggplot(agg_data_pie, aes(x = "", y = Percentage_only_exports_fob, fill = {{show_group}})) +
      geom_bar(width = 1, stat = "identity") +
      coord_polar("y", start = 0) +
      theme_void()
  )

  # print(p1)
  # print(p2)
  print(plot_grid(p1, p2, legend, ncol = 3, labels = c("Only Exports fob", "Only Exports qty"), 
                  rel_widths = c(1, 1, 0.4)))
  print(plot_grid(p3, p4, legend, ncol = 3, labels = c("ReExports fob", "ReExports qty"), 
                  rel_widths = c(1, 1, 0.4)))
  # print(p3)
  # print(p4)
}
unitgroup <- exports_raw  %>% plotv2(show_group=continent,month,continent)
print(unique(unitgroup$show_group))
unitgroup %>% pie_plot(continent)

```

top n contries,
The FOB curve for Antarctica seems strange, with significant fluctuations, possibly due to its unique geographic location.

```{r}
library(treemapify)
library(highcharter)
# tree_map <- exports_raw %>%
#   group_by(hs2_group, hs2) %>%
#   summarise(total = sum(total_exports_nzd_fob)) %>%
#   ggplot(aes(area = total, fill = hs2_group,
#            label = hs2, subgroup = hs2_group)) +
#   geom_treemap() +
#   geom_treemap_subgroup_border() +
#   geom_treemap_text(colour = "white") +
#   scale_fill_viridis_d()
# print(tree_map)
# data_to_hierarchical(
#   exports_raw,
#   group_vars = c("hs2_group", "hs2"),
#   size_var = "total_exports_nzd_fob",
# )
# exports_raw %>%
#   group_by(hs2_group, hs2) %>%
#   summarise(total = sum(total_exports_nzd_fob)) %>%
#   plot_ly(
#       type = "treemap",
#       labels = ~hs4,
#       parents = ~hs2_group,
#       values = ~total
#       # branchvalues = "total",
#       # hovertemplate = '<b>%{label}</b><br>价值: $%{value:,.2f}<br>章节: %{parent}<extra></extra>'
#     ) %>%
#       layout(
#         title = "出口商品结构",
#         margin = list(l = 0, r = 0, b = 0, t = 50)
#       )
# ggplotly(tree_map,tooltip = c("label", "area"))
```


## Continent-Product Bicluster Heatmap

```{r}
library(ComplexHeatmap)
matrix_data <- exports_raw %>%
  group_by(continent, hs2_group) %>%
  summarise(total = log10(sum(total_exports_nzd_fob))) %>%
  spread(hs2_group, total, fill=0) %>%
  column_to_rownames("continent")

Heatmap(matrix_data, 
        name = "log10(Export Value)",
        col = viridisLite::viridis(100),
        row_km = 4,  # 国家聚类
        column_km = 4)  # 产品聚类
```

Ships' Bunkering、Ships' Stores 和 Passengers' Effects 这三类出口的比例相似，符合预期，均与旅游相关。相比之下，Antarctica 和目的不明的出口量较少。

## three demension analysis(total_value,avg_price,product_diversity)

```{r}
country_stats <- exports_raw %>%
  group_by(continent) %>%
  summarise(
    total_value = sum(total_exports_nzd_fob),
    avg_price = mean(total_exports_nzd_fob/total_exports_qty, na.rm=T),
    product_diversity = n_distinct(harmonised_system_code)
  ) 

ggplot(country_stats, aes(x=log10(total_value), y=log10(avg_price),
                         size=product_diversity, color=continent)) +
  geom_point(alpha=0.7) +
  ggrepel::geom_text_repel(aes(label=continent), size=3, max.overlaps = 15) +
  scale_color_brewer(palette = "Set2") +
  labs(title = "Market Dynamics Analysis",
       subtitle = "Size=Product Diversity, Color=Continent")
```

## group_by hs2_group


```{r}
exports_raw %>% plotv2(show_group=hs2_group,month,hs2_group)
```

```{r}
# # 洛伦兹曲线与基尼系数
# country_share <- exports_raw %>%
#   group_by(country) %>%
#   summarise(total = sum(total_exports_nzd_fob)) %>%
#   arrange(desc(total)) %>%
#   mutate(
#     cum_share = cumsum(total)/sum(total),
#     rank = row_number()
#   )
# 
# n_countries <- nrow(country_share)
# 
# gini <- country_share %>%
#   summarise(gini = 1 - 2 * sum(cum_share)/n_countries)
# 
# 
# ggplot(country_share) +
#   geom_line(aes(x = rank/n_countries, y = cum_share), color="steelblue") +
#   geom_abline(slope = 1, linetype="dashed") +
#   annotate("text", x=0.7, y=0.3, label=paste0("Gini Index: ", round(gini$gini,3))) +
#   labs(title = "Export Market Concentration (Lorenz Curve)",
#        x = "Cumulative Percentage of Countries",
#        y = "Cumulative Share of Exports")
```


## diversity analysis 

```{r}
# 多样性分析可视化
library(vegan)
library(rnaturalearth)
# 2. 产品多样性分析（按国家）
product_diversity <- exports_raw %>%
  group_by(iso3, hs2_group) %>%
  summarise(total_value = sum(total_exports_nzd_fob)) %>%
  group_by(iso3) %>%
  mutate(
    shannon = diversity(total_value, index = "shannon"),
    # 计算Simpson指数并计算市场集中度（1 - Simpson指数）
    simpson_index = diversity(total_value, index = "simpson"),
    market_concentration = 1 - simpson_index
  )

# 产品多样性地图
world_shannon <- ne_countries(returnclass = "sf") %>%
  left_join(product_diversity, by = c("iso_a3" = "iso3"))

plot_geo(product_diversity) %>%
  add_trace(
    z = ~shannon,
    locations = ~iso3,
    colorscale = "Viridis"
  ) %>%
  layout(title = "Shannon Index")

# 3. 多样性-集中度关系分析
ggplot(product_diversity, aes(x = shannon, y = market_concentration)) +
  geom_point(aes(size = total_value), alpha = 0.7) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  scale_size_continuous(name = "total export fob（log）", 
                       trans = "log10",
                       breaks = c(1e6, 1e7, 1e8)) +
  scale_color_brewer(palette = "Set2") +
  labs(title = "Relationship between diversity and market concentration",
       x = "product diversity indexs", 
       y = "market concentration") +
  theme_bw()

# 4. 多样性时间趋势分析
diversity_trend <- exports_raw %>%
  mutate(quarter = quarter(month)) %>%
  group_by(quarter, country) %>%
  summarise(total_value = sum(total_exports_nzd_fob)) %>%
  group_by(quarter) %>%
  mutate(
    shannon = diversity(total_value, index = "shannon")
  ) %>%
  distinct(quarter, shannon)

ggplot(diversity_trend, aes(x = quarter, y = shannon)) +
  geom_line(color = "#2c7fb8", size = 1.5) +
  geom_point(color = "#2c7fb8", size = 3) +
  labs(title = "Market Diversity Quarterly Trends",
       x = "quarter", y = "Shannon Diversity Index") +
  scale_x_continuous(breaks = 1:4, labels = paste0("Q", 1:4)) +
  theme_minimal()
```